---
title: "Visualization"
author: "Madeline Sharp"
date: "2026-10-01"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

Now we have everything!

```{r}
weather_df

#don't put in code
#weather_df |> view()
```

Let's make a scatterplot!

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

I always like the dataframe first.

```{r}
ggp_temp_scatterplot =
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_temp_scatterplot
```

Let's make this a bit fancier...

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  # makes points abit transparent
  geom_point(alpha = .25) +
  # adds a smooth line through the middle. 
  # se = FALSE gets rid of error bars. Jeff prefers no error bars
  geom_smooth(se = FALSE)
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  # different aesthetic definition
  geom_point(aes(color = name), alpha = .25) +
  geom_smooth(se = FALSE)
```

The aesthetics are up to you

```{r}
# Order: dataset -> mappings -> what you want to see
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

Show faceting

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  # how you want columns to be separated ". ~ name", i.e., by name
  # classic interface
  facet_grid(. ~ name)
```

Changing orientation of facet

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  facet_grid(name ~.)
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  # new interface. Does the same thing as the classic one.
  facet_grid(cols = vars(name))
```

Let's look at something else.

```{r}
weather_df |>
  # w/o specifications R creates scales on its own, e.g. with dates
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

make a plot of central park tmax v tmin only, and convert temperatures to 
fahrenheit.

```{r}
weather_df |>
  # clean data outside ggplot before plotting
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax = tmax * 9/5 + 32,
    tmix = tmin * 9/5 + 32
  ) |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() 
```


what's a hex plot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

## Univariate plots

```{r}
weather_df |>
  ggplot(aes(x = tmax)) +
  geom_histogram()
```

```{r}
weather_df |>
  # Jeff does not like this.
  ggplot(aes(x = tmax, color = name)) +
  geom_histogram()
```

```{r}
weather_df |>
  # Use fill instead
  ggplot(aes(x = tmax, fill = name)) +
  # Jeff does not like this
  geom_histogram(position = "dodge")
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  # Use this instead
  facet_grid(. ~ name)
```

Density plots are great!! (Density plots are smoothed out histograms)

```{r}
weather_df |>
  ggplot(aes(x = tmax, color = name)) +
  geom_density()
```

```{r}
weather_df |>
  # Fill covers some data here. Not good.
  ggplot(aes(x = tmax, fill = name)) +
  geom_density()
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  # Use alpha with fill if you want to make the colours more transparent.
  # Otherwise will cover other data. 
  # This is better.
  geom_density(alpha = .3)
```

Boxplots

```{r}
# Histograms are better for bimodal dists.
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin - violin plots super popular circa 2018. Same with ridge plots

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  # is very useful when you are looking at distributions in many areas
  # e.g. distribution over many US states
  geom_violin()
```

Ridge plots...

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = name)) +
  # is very useful when you are looking at distributions in many areas
  # e.g. distribution over many US states
  # more useful than others for this
  geom_density_ridges()
```

Remember what aesthetic mapping make sense!

## Save some of the plots

```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  facet_grid(. ~ name)

ggp_weather

ggsave("ggp_weather.pdf", ggp_weather)

# if you find yourself saving a lot of images, save them into a subfolder
```

```{r fid.width = 12, fig.height = 12}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
# figure out what this does
#knitr::opts_chunk$set(
#  fig.width = 6,
#  fig.asp = .6,
#  out.width = "90%"
#)
```